---
title: "power"
author: "Nathan Young"
date: "11/9/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#libraries 
library(tidyverse)
library(pwr)

```

#pwr

pwr is an r package that can help you calculate power (or other parameters) for a number of different tests:

t-test
correlation
proportional tests
one-way anova
chi-squared tests
multiple regression models

##T-tests

```{r}
#power analysis for t-tests using the pwr package

#leave one term out to calculate it. Fill in the rest.
pwr.t.test(n = , d = , sig.level = , power = , type = )

#see examples below

```

```{r}
#What sample size do we need to detect at minimum a d = .3, at an alpha =.05, and power at least .80?

pwr.t.test(d = .3, sig.level = .05, power = .8, type = "two.sample")

#Switch "two.sample" to "paired" for within subjects

```

```{r}
#view power curve
powerout <- pwr.t.test(d = .3, sig.level = .05, power = .8, type = "two.sample")

plot(powerout) + theme_bw()
```


```{r}
#What power would/do we have?

pwr.t.test(n = 150, d = .3, sig.level = .05, type = "two.sample")
```


```{r}
#Sensitivity  - calculate critical effect size

#What is the minimum effect size we can reliably detect given that we collect 100 participants, at an alpha of .05, with at least power of .80?

pwr.t.test(n = 100, power = .8, sig.level = .05, type = "two.sample")

```

##correlation

```{r}
#leave one out, fill in the rest to calculate the left out parameter
pwr.r.test(n =, r =, sig.level = , power = )
```

```{r}
#what sample size do we need for an r = .3, alpha of .05, and power at least of .80?
pwr.r.test(r = .3, sig.level = .05, power = .8)
```

#SuperPower
https://aaroncaldwell.us/SuperpowerBook/

online Shiny apps (see https://arcstats.io/shiny/anova-exact/ and https://arcstats.io/shiny/anova-power/).
```{r}
#power analysis with superpower
library(Superpower)

```

Superpower allows researchers to perform simulation-based power analyses without having extensive programming knowledge.

We can do an example from the SuperPower paper: Lakens & Caldwell (2019)

##2 between subjects conditions
```{r}
#first we specify a design and the expected distributions of the conditions

design_result <- ANOVA_design(
  design = "2b", 
  n = 100,
  mu = c(.8, 0), 
  sd = 2,
  labelnames = c("condition", 
                 "cheerful", 
                 "sad"),
  plot = TRUE)

#we have 2 between subjects conditions: design = ("2b")
#we base our analysis on a sample of 80 participants per condition: n = 80
#the means of those conditions are 1 and 0: mu = c(1,0)
#the standard deviation across those conditions is 2: sd = 2

#see visualization created below once this block of code is run

#for reference cohen's d = (m1 - m2)/sd
#this example: (0-1)/2 = -.5
```

```{r}
mean1 <-.8
mean2 <-0
sd_ <-2

cohensd <- (mean1 - mean2)/sd_

cohensd
```

```{r}

condition <- rep(letters[1], each = 1000)
score <- rnorm(n = 1000, mean = mean1, sd = sd_)

A <- cbind(condition, score)

condition <- rep(letters[2], each = 1000)
score <- rnorm(n = 1000, mean = mean2, sd = sd_)

B <- cbind(condition, score)

data <- rbind(A,B)

data <- as.tibble(data)

data$score <- as.numeric(data$score)

data %>%
  ggplot(aes(x = score, fill = condition)) +
  geom_density(alpha = .3) + 
  geom_vline(aes(xintercept=0, na.rm=T),   # Ignore NA values for mean
               color="red", linetype="dashed", size=1) +
  geom_vline(aes(xintercept = 1, na.rm=T),
             color="blue", linetype="dashed", size = 1) +
  ggtitle("Representation of a Cohen's d of .4")
  
```



```{r}
#next we simulate a data set that has the properties we specified above

#there are 2 ways to do that with the SuperPower package

#1 - ANOVA_power - this function simulates data sets repeatedly based on the specified parameters and caculates the percentage of statistically significant results


#2 - ANOVA_exact - simulates a data set that has EXACTLY the desired properties, performs an ANOVA, and uses the ANOVA results to compute power
```

```{r}
#1 - ANOVA_power: this will take a few minutes (depends on your computer)
#the more nsims the closer the result will be to #2

design_sim <- ANOVA_power(design_result, 
                   nsims = 5000)
```

```{r}
#2 - ANOVA_exact - **recommended**

design_exact <- ANOVA_exact(design_result)
```

##3 between subjects conditions

```{r}
design_result_1 <- ANOVA_design(
  design = "3b", 
  n = 80,
  mu = c(1, 0.5, 0), 
  sd = 2,
  labelnames = c("condition",
                 "cheerful",
                 "neutral", 
                 "sad"))

#effect size reference: Cohen's f = standard deviation of the means / standard deviation

#this example: .5/2 = .25

#to patial eta squared (n2p): sqrt((.25^2) / ((.25^2)+1)) = n2p 
```

```{r}
mean1 <-1
mean2 <-.5
mean3 <- 0
sd_ <-2

cohensf <- (sd(c(mean1,mean2,mean3))/sd_)

cohensf
```

```{r}

condition <- rep(letters[1], each = 1000)
score <- rnorm(n = 1000, mean = mean1, sd = sd_)

A <- cbind(condition, score)

condition <- rep(letters[2], each = 1000)
score <- rnorm(n = 1000, mean = mean2, sd = sd_)

B <- cbind(condition, score)

condition <- rep(letters[3], each = 1000)
score <- rnorm(n = 1000, mean = mean3, sd = sd_)

C <- cbind(condition, score)

data <- rbind(A,B,C)

data <- as.tibble(data)

data$score <- as.numeric(data$score)

data %>%
  ggplot(aes(x = score, fill = condition)) +
  geom_density(alpha = .3) + 
  geom_vline(aes(xintercept=1, na.rm=T),   # Ignore NA values for mean
               color="red", linetype="dashed", size=1) +
  geom_vline(aes(xintercept = .5, na.rm=T),
             color="green", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 0, na.rm=T),
             color="blue", linetype="dashed", size = 1) +
  ggtitle("Representation of a Cohen's f of .25")
  
```

```{r}
#2 - ANOVA_exact - **recommended**

design_exact <- ANOVA_exact(design_result_1)
```

##within subjects design

```{r}
design_within <- ANOVA_design(
  design = "3w", 
  n = 80, 
  mu = c(1, 0.5, 0),
  sd = 2, 
  r = 0.5,
  labelnames = c("condition",
                 "cheerful",
                 "neutral", 
                 "sad"))

#new parameters - r

#r = correlation between repeated measures

#effect size reference - Cohen's f 
```

```{r}
#run a simulation based on the design
power_within = ANOVA_power(design_within, nsims = 100000)

#OR

exact_within = ANOVA_exact(design_within)
```
##Power for interactions

```{r}
design_result_cross <- ANOVA_design(
  design = "2b*2b", 
  n = 120,
  mu = c(1, 0, 0, 0), 
  #c(a1_b1, a1_b2, a2_b1, a2_b2) - order of condition means for a 2 x 2
  sd = 2,
  labelnames = c("condition",
                 "cheerful", "sad",
                 "voice",
                 "human", "robot"))

#effect size reference - Cohen's f for interaction effects - Box 3 Lakens & Caldwell (2019)

#see page 8 box 3 in Lakens & Caldwell (2019) for important considerations regarding crossover vs non-crossover interactions

#Also this talk by Jeff Rouder - https://www.youtube.com/watch?v=PzHcwS3xbZ8
```

```{r}
mean1 <-1
mean2 <-0
mean3 <-0
mean4 <-0
sd_ <-2

gm <- ((mean1 + mean2 + mean3 + mean4)/4)
mm1 <- mean(c(mean1, mean2)) - gm
mm2 <- mean(c(mean3, mean4)) - gm
mm3 <- mean(c(mean2, mean4)) - gm
mm4 <- mean(c(mean1, mean3)) - gm


cohensf <- ((sqrt(((((mean1 - (gm + mm1 + mm4))^2) +
                ((mean2 - (gm + mm1 + mm3))^2) +
                ((mean3 - (gm + mm2 + mm4))^2) +
                ((mean4 - (gm + mm2 + mm3))^2))/4))/2))

cohensf
```

```{r}

conditiona <- rep(letters[1], each = 1000)
conditionb <- rep(letters[3], each = 1000)
score <- rnorm(n = 1000, mean = mean1, sd = sd_)

A <- cbind(conditiona, conditionb, score)

conditiona <- rep(letters[1], each = 1000)
conditionb <- rep(letters[4], each = 1000)
score <- rnorm(n = 1000, mean = mean2, sd = sd_)

B <- cbind(conditiona, conditionb, score)

conditiona <- rep(letters[2], each = 1000)
conditionb <- rep(letters[3], each = 1000)
score <- rnorm(n = 1000, mean = mean3, sd = sd_)

C <- cbind(conditiona, conditionb, score)

conditiona <- rep(letters[2], each = 1000)
conditionb <- rep(letters[4], each = 1000)
score <- rnorm(n = 1000, mean = mean4, sd = sd_)

D <- cbind(conditiona, conditionb, score)

data <- rbind(A,B,C,D)

data <- as.tibble(data)

data$score <- as.numeric(data$score)

data %>%
  ggplot(aes(x = score, fill = conditiona)) +
  geom_density(alpha = .3) + 
  geom_vline(aes(xintercept=1, na.rm=T),   # Ignore NA values for mean
               color="yellow", linetype="dashed", size=1) +
  geom_vline(aes(xintercept = 0, na.rm=T),
             color="black", linetype="dashed", size = 1) +
  ggtitle("Representation of a Cohen's f of .25") + 
  facet_grid(~conditionb)
  
```

```{r}
exact_cross = ANOVA_exact(design_result_cross)
```

##power curves

```{r}
design_result_eg <- ANOVA_design(
  design = "2b*2b", 
  n = 50,
  mu = c(30, 35, 30, 31), 
  sd = 5,
  labelnames = c("AgeGroup",
                 "Young", "Old",
                 "Emotion",
                 "Neutral", "Anger"))


plot_power(design_result_eg,
           min_n = 10, 
           max_n = 100,
           desired_power = 80, 
           plot = TRUE)
```

